---
title: "206 Assignment 4"
author: "Sidney Gerst"
date: "November 13, 2018"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load Packages and read in files

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(pwr)
library(knitr)
library(kableExtra)
library(plotly)
library(ggrepel)
library(effsize)
library(vcdExtra)
library(dplyr)
library(car)

lobster_pressure <- read_csv("lobster_traps.csv")

lobster_size <- read_csv("lobster_size_abundance.csv")



```

Clean up Data before converting to tidy

```{r, echo=FALSE, message=FALSE}
size_clean <- lobster_size %>%
  select(YEAR, SITE, SIZE, COUNT) %>%
  filter(COUNT != "0")

size_clean

pressure_clean <- lobster_pressure %>%
  select(YEAR, SITE, TRAPS) 
  
pressure_clean



```


Part 1: Lobster abundance and fishing pressure
- Exploratory data analysis
    - Scatter plots with 5 locations with counts and fishing pressure (traps) for 2012-2017 (separately)
    - Column graphs with total count/traps by site
    - Finalized line graphs for abundance and pressure
    
    
```{r, echo=FALSE, message=FALSE}

## Make scatter plot for annual abundance by site (2012-2017)
abundance <- size_clean %>% 
  select(YEAR, SITE, COUNT) %>% 
  group_by(SITE, YEAR) %>% 
  summarize (
      total_count = round(sum(COUNT), digits = 1)
    ) %>% 
  ggplot(aes(x=YEAR, y = total_count))+
  geom_point() +
  facet_wrap(~SITE, scale = "free") +
  xlab("Year") +
  ylab("Total Count")+
  theme_classic()+
  ggtitle("Total Lobster Abundance (2012-2017) - TH, EM, SG")
  

abundance


## Make column graph of total abundance by site (2012-2017)
abundance_col <- ggplot(size_clean, aes(x = YEAR, y = COUNT))+
  geom_col()+
  facet_wrap(~SITE, scale = "free")+
  theme_classic()+
  ggtitle("Total Lobster Abundance (2012-2017) - TH, EM, SG")

abundance_col


## Make line graph of total abundance within each site (2012-2017)
abundance_line <- as.data.frame(size_clean) %>% 
  expand.dft(freq="COUNT") %>% 
  count(YEAR, SITE) %>% 
  ggplot(aes(x=YEAR, y = n))+
  geom_line(aes(color=SITE)) +
  xlab("Year") +
  ylab("Total Count")+
  ggtitle("Lobster Abundance at LTER Sites in the Santa Barbara Channel - TH, EM, SG")+
  theme_classic()
  

abundance_line


```


```{r, echo=FALSE, message=FALSE}

# Make a scatter plot for annual lobster pressure for each site (2012-2017)
pressure <- pressure_clean %>% 
  select(YEAR, SITE, TRAPS) %>%
  filter(SITE == "AQUE"| SITE == "NAPL" | SITE == "MOHK"| SITE == "IVEE"| SITE == "CARP") %>% 
  group_by(SITE, YEAR) %>% 
  summarize (
      total_traps = round(sum(TRAPS), digits = 1)
    ) %>% 
  ggplot(aes(x=YEAR, y = total_traps))+
  geom_point() +
  facet_wrap(~SITE, scale = "free") +
  xlab("Year") +
  ylab("Total Pressure")+
  ggtitle("Lobster Abundance (2012-2017) - TH, EM, SG")+
  theme_classic()

pressure


# Make final pressure data frame (with just 5 sites)
pressure_final <- pressure_clean %>% 
  select(YEAR, SITE, TRAPS) %>%
  filter(SITE == "AQUE"| SITE == "NAPL" | SITE == "MOHK"| SITE == "IVEE"| SITE == "CARP")
 

# Make a column graph for total pressure for each site (2012-2017)
pressure_col <- ggplot(pressure_final, aes(x = YEAR, y = TRAPS))+
  geom_col()+
  facet_wrap(~SITE, scale = "free")+
  theme_classic()+
  ggtitle("Total Lobster Fishing Pressure (2012-2017) - TH, EM, SG")

pressure_col


# Make line graph of total pressure within each site (2012-2017)
pressure_line <- as.data.frame(pressure_final) %>% 
  expand.dft(freq="TRAPS") %>% 
  count(YEAR, SITE) %>% 
  ggplot(aes(x=YEAR, y = n))+
  geom_line(aes(color=SITE)) +
  xlab("Year") +
  ylab("Total Pressure")+
  ggtitle("Lobster Fishing Pressure at LTER Sites in Santa Barbara Channel - TH, EM, SG")+
  theme_classic()

pressure_line



```

Discuss trends in lobster abundance (counts) and fishing pressure (buoys):
- In each of the five sites, the total lobster abundance is greater in 2017 than it was in 2012. However, counts vary greatly over the 5 years. At the AQUE site, counts were much greater over the course of all 5 years than at the other sites. CARP had a dramatic increase in 2017, whereas prior to that the counts were very low. AQUE, CARP, and IVEE all ended 2017 with counts higher than the previous year, while the other two sites, MOHK and NAPL, experienced decline in their final year. An analysis of lobster pressure may partially explain trends in lobster abundance. Within the AQUE, MOHK, and CARP sites there was a decline in lobster fishing pressure between 2016 and 2017. This could have caused the significant increase in lobster abundance seen at the CARP site within the same time frame. In general, lobster fishing pressure decreased between 2012 and 2017 at CARP, MOH, and AQUE. NAPL and IVEE are Marine Protected Areas (MPA's) and so have zero fishing pressure. 



Part 2. Compare mean lobster sizes (by carapice length (mm)) by site in 2017.

```{r}

# Convert abundance data frame into tidy format

abundance_tidy <- as.data.frame(size_clean) %>% 
  expand.dft(freq="COUNT")

abundance_tidy

#Create data frame for our 5 locations in 2017, selecting size.
sizes_2017 <- abundance_tidy %>% 
  filter(YEAR == 2017,
         SITE == "AQUE"|SITE == "NAPL"|SITE == "MOHK"|SITE == "IVEE"|SITE == "CARP",
         SIZE != "-99999")
View(sizes_2017)

# Basic data exploration

size_hist <- ggplot(sizes_2017, aes(x = SIZE)) +
  geom_histogram(bins = 10) +
  facet_wrap(~ SITE, scale = "free")

size_hist
  
size_qq <- ggplot(sizes_2017, aes(sample = SIZE)) +
  geom_qq(bins = 10) +
  facet_wrap(~ SITE, scale = "free")

size_qq

#We see the data is approximately normally distributed at every site in 2017.

box_size <- ggplot(sizes_2017, aes(x = SITE, y = SIZE)) +
  geom_boxplot(width = .4, aes(color = SITE))
  

box_size

#Conduct a Levene's Test to see if equal variances, so that we can use an ANOVA.

variances <- sizes_2017 %>% 
  group_by(SITE) %>% 
  summarize(
    mean = mean(SIZE),
    sd = sd(SIZE),
    variance = var(SIZE)
  )

variances

levene_var <- leveneTest(SIZE ~ SITE, data = sizes_2017)
levene_var

#H0:Variances are equal
#HA: Variances are not equal

#With such a small p-value, we choose to reject the null. There is a significant difference in variances between sites.


#Conduct a one-way ANOVA.
#H0: All means are equal
#HA: At least two means are not equal

size_anova <- aov(SIZE ~ SITE, data = sizes_2017)

size_sum <- summary(size_anova)

size_sum ##ANOVA result: weak significance (p=0.0085)

#Tukey's post hoc test - Which means differ?

size_tukey <- TukeyHSD(size_anova)
size_tukey
#Significant difference in mean lobster carapice sizes between NAPL-IVEE (p=0.0037)


```
There was not a significant difference in the mean lengths of lobsters at the LTER sites in 2017, with the exception of Naples and Isla Vista, which had an absolute difference of 4.78mm. 


Part 3. Changes in lobster size at MPA and non-MPA sites (comparing only 2012 and 2017 sizes)
  - 1.  Exploratory data analysis
        - Summary data table with size means within MPA sites during 2012 and 2017
        - Summary data table with size within non-MPA sites during 2012 and 2017
        - Histograms and QQplots for MPA and Non-MPA sites (2012-2017)
  - 2. Hypothesis test for size change in MPA sites during 2012 and 2017
        - Null Hypothesis -- there is no significant difference in carapace length (2012 and 2017)
        - Alternative Hypothesis -- there IS a significant difference in carapace length (2012 and             2017)
  - 3. Hypothesis test for size change in non-MPA sites during 2012 and 2017
        - Null Hypothesis -- there is no significant difference in carapace length between 2012 and            2017
        - Alternative Hypothesis -- there IS a significant difference in carapace length (2012-2017)
  - 4. Hypothesis test to compare carapace length between MPA sites and non-MPA sites during 2012           (and separately 2017)
        - Null Hypothesis -- there is no significant difference in carapace length of lobsters in MPA          sites and non-MPA sites
        - Alternative Hypothesis -- there IS no significant difference in carapace length of lobsters          in MPA sites and non-MPA sites
      
```{r}

################################
### 1. EXPLORATORY DATA ANALYSIS
################################


## Set up data frame that assigns MPA or non-MPA designation
MPA_sites <- abundance_tidy %>% 
  filter(YEAR == "2012" | YEAR == "2017") %>% 
  mutate(
    STATUS = case_when(
      SITE == "IVEE" ~ "MPA", 
      SITE == "NAPL" ~ "MPA",
      SITE == "AQUE" ~ "Non-MPA",
      SITE == "MOHK" ~ "Non-MPA",
      SITE == "CARP" ~ "Non-MPA")) %>% 
  select(YEAR, SITE, SIZE, STATUS)

## Create summary data table for MPA sites with mean, median, max, SD, sample size for SIZE at each site for 2012 and 2017  (IVEE, NAPL)
MPA_size_table <- MPA_sites %>% 
  filter(STATUS == "MPA") %>% 
  group_by(YEAR) %>% 
  summarize(
    mean_size = round(mean(SIZE), 2), 
    median_size = round(median(SIZE), 2),
    max_size = round(max(SIZE), 2),
    sd_size = round(sd(SIZE), 2),
    length(SIZE)
  )

MPA_size_table

## Create summary data table for NON-MPA sites with mean, median, max, SD, sample size for SIZE at each site for 2012 and 2017 (AQUE, MOHK, CARP)
Non_MPA_size_table <- MPA_sites %>% 
  filter(STATUS == "Non-MPA") %>% 
  group_by(YEAR) %>% 
  summarize(
    mean_size = round(mean(SIZE), 2), 
    median_size = round(median(SIZE), 2),
    max_size = round(max(SIZE), 2),
    sd_size = round(sd(SIZE), 2),
    length(SIZE)
  )

Non_MPA_size_table


#########################
# MPA Sites Data Analysis
#########################


## Create finalized histogram for MPA sites during 2012
MPA_hist_2012 <- MPA_sites %>% 
  filter(STATUS == "MPA") %>%
  filter(YEAR == "2012") %>% 
  ggplot(aes(x=SIZE))+
  geom_histogram(bins = 11, boundary= 0, colour="white", fill= "deepskyblue4")+
  facet_wrap(~SITE, scale = "free")+
  xlab("Carapace Length(mm)")+
  ylab("Abundance")+
  ggtitle("Lobster Sizes at MPA Sites during 2012 - TH, EM, SG")+
  scale_y_continuous(expand = c(0,0))+
  scale_x_continuous(expand = c(0,0))+
  theme_classic()

MPA_hist_2012

## Create qqplot for MPA sites during 2012

MPA_qq_2012 <- MPA_sites %>% 
  filter(STATUS == "MPA") %>%
  filter(YEAR == "2012") %>% 
  ggplot(aes(sample=SIZE))+
  geom_qq()+
  facet_wrap(~SITE, scale = "free")+
  ylab("Carapace Length(mm)")+
  ggtitle("Lobster Sizes at MPA Sites during 2012 - TH, EM, SG")+
  theme_classic()

MPA_qq_2012


## Create finalzied histogram for MPA sites during 2017
MPA_hist_2017 <- MPA_sites %>% 
  filter(STATUS == "MPA") %>%
  filter(YEAR == "2017") %>% 
  ggplot(aes(x=SIZE))+
  geom_histogram(bins = 11, boundary= 0, colour="white", fill= "deepskyblue4")+
  facet_wrap(~SITE, scale = "free")+
  xlab("Carapace Length(mm)")+
  ylab("Abundance")+
  ggtitle("Lobster Sizes at MPA Sites during 2017 - TH, EM, SG")+
  scale_y_continuous(expand = c(0,0))+
  scale_x_continuous(expand = c(0,0))+
  theme_classic()
      
MPA_hist_2017


## Create qqplot for MPA sites during 2017
MPA_qq_2017 <- MPA_sites %>% 
  filter(STATUS == "MPA") %>%
  filter(YEAR == "2017") %>% 
  ggplot(aes(sample=SIZE))+
  geom_qq()+
  facet_wrap(~SITE, scale = "free")+
  ylab("Carapace Length(mm)")+
  ggtitle("Lobster Sizes at MPA Sites during 2017 - TH, EM, SG")+
  theme_classic()

MPA_qq_2017


## Create finalized histogram for MPA sites grouped by YEAR
MPA_hist_year <- MPA_sites %>% 
  filter(STATUS == "MPA") %>%
  ggplot(aes(x=SIZE))+
  geom_histogram(bins = 13, boundary= 0, colour="white", fill= "deepskyblue4")+
  facet_wrap(~YEAR, scale = "free")+
  xlab("Carapace Length(mm)")+
  ylab("Abundance")+
  ggtitle("Lobster Sizes at MPA Sites during 2012 and 2017 - TH, EM, SG")+
  scale_y_continuous(expand = c(0,0))+
  scale_x_continuous(expand = c(0,0))+
  theme_classic()

MPA_hist_year

## Create qqplot for MPA sites grouped by YEAR
MPA_qq_year <- MPA_sites %>% 
  filter(STATUS == "MPA") %>%
  ggplot(aes(sample=SIZE))+
  geom_qq()+
  facet_wrap(~YEAR, scale = "free")+
  ylab("Carapace Length(mm)")+
  ggtitle("Lobster Sizes at MPA Sites during 2012 and 2017 - TH, EM, SG")+
  theme_classic()

MPA_qq_year


##############################
# NON - MPA Sites Data Analysis
##############################

## Create finalized histogram for NON - MPA sites during 2012
Non_MPA_hist_2012 <- MPA_sites %>% 
  filter(STATUS == "Non-MPA") %>%
  filter(YEAR == "2012") %>% 
  ggplot(aes(x=SIZE))+
  geom_histogram(bins = 11, boundary= 0, colour="white", fill= "deepskyblue4")+
  facet_wrap(~SITE, scale = "free")+
  xlab("Carapace Length(mm)")+
  ylab("Abundance")+
  ggtitle("Lobster Sizes at Non-MPA Sites during 2012 - TH, EM, SG")+
  scale_y_continuous(expand = c(0,0))+
  scale_x_continuous(expand = c(0,0))+
  theme_classic()

Non_MPA_hist_2012

## Create qqplot for MPA sites during 2012

Non_MPA_qq_2012 <- MPA_sites %>% 
  filter(STATUS == "Non-MPA") %>%
  filter(YEAR == "2012") %>% 
  ggplot(aes(sample=SIZE))+
  geom_qq()+
  facet_wrap(~SITE, scale = "free")+
  ylab("Carapace Length(mm)")+
  ggtitle("Lobster Sizes at Non-MPA Sites during 2012 - TH, EM, SG")+
  theme_classic()

Non_MPA_qq_2012


## Create finalzied histogram for NON - MPA sites during 2017
Non_MPA_hist_2017 <- MPA_sites %>% 
  filter(STATUS == "Non-MPA") %>%
  filter(YEAR == "2017") %>% 
  ggplot(aes(x=SIZE))+
  geom_histogram(bins = 11, boundary= 0, colour="white", fill= "deepskyblue4")+
  facet_wrap(~SITE, scale = "free")+
  xlab("Carapace Length(mm)")+
  ylab("Abundance")+
  ggtitle("Lobster Sizes at Non-MPA Sites during 2017 - TH, EM, SG")+
  scale_y_continuous(expand = c(0,0))+
  scale_x_continuous(expand = c(0,0))+
  theme_classic()
      
Non_MPA_hist_2017


## Create qqplot for NON - MPA sites during 2017
Non_MPA_qq_2017 <- MPA_sites %>% 
  filter(STATUS == "Non-MPA") %>%
  filter(YEAR == "2017") %>% 
  ggplot(aes(sample=SIZE))+
  geom_qq()+
  facet_wrap(~SITE, scale = "free")+
  ylab("Carapace Length(mm)")+
  ggtitle("Lobster Sizes at Non-MPA Sites during 2017 - TH, EM, SG")+
  theme_classic()

Non_MPA_qq_2017


## Create finalized histogram for NON - MPA sites grouped by YEAR
Non_MPA_hist_year <- MPA_sites %>% 
  filter(STATUS == "Non-MPA") %>%
  ggplot(aes(x=SIZE))+
  geom_histogram(bins = 13, boundary= 0, colour="white", fill= "deepskyblue4")+
  facet_wrap(~YEAR, scale = "free")+
  xlab("Carapace Length(mm)")+
  ylab("Abundance")+
  ggtitle("Lobster Sizes at Non-MPA Sites during 2012 and 2017 - TH, EM, SG")+
  scale_y_continuous(expand = c(0,0))+
  scale_x_continuous(expand = c(0,0))+
  theme_classic()

Non_MPA_hist_year

## Create qqplot for NON - MPA sites grouped by YEAR
Non_MPA_qq_year <- MPA_sites %>% 
  filter(STATUS == "Non-MPA") %>%
  ggplot(aes(sample=SIZE))+
  geom_qq()+
  facet_wrap(~YEAR, scale = "free")+
  ylab("Carapace Length(mm)")+
  ggtitle("Lobster Sizes at Non-MPA Sites during 2012 and 2017 - TH, EM, SG")+
  theme_classic()

Non_MPA_qq_year



##########################
### 2. HYPOTHESIS TESTING 
##########################

#Hypothesis test for size change in MPA sites during 2012 and 2017
  ## Null Hypothesis -- there is no significant difference in carapace length (2012 and 2017)
  ## Alternative Hypothesis -- there IS a significant difference in carapace length (2012 and             2017)
 
```


Part 4. Proportions of “legal” lobsters at the 5 sites in 2017

The legal minimum carapace size for lobster is 82.6 mm. What proportion of observed lobsters at each site are above the legal minimum? Does that proportion differ significantly across the 5 sites?

Steps for chi-square test:
  -Create table of proportions for carapace size in 2017 for the 5 sites
  -Chi-square test

```{r echo = false}
#table of proportions using the df: sizes_2017
sizes_17 <- sizes_2017 %>% 
  filter( YEAR == "2017") %>% 
  mutate(PROTECTED = case_when(
    SITE == "IVEE" ~ "MPA",
    SITE == "NAPL" ~ "MPA",
    SITE == "AQUE" ~ "Non-MPA",
    SITE == "CARP" ~ "Non-MPA",
    SITE == "MOHK" ~ "Non-MPA"))

sizes_17

sizes_17_summary <- sizes_2017 %>% 
  filter(YEAR == "2017") %>% 
  mutate(PROTECTED = case_when(
    SITE == "IVEE" ~ "MPA",
    SITE == "NAPL" ~ "MPA",
    SITE == "AQUE" ~ "Non-MPA",
    SITE == "CARP" ~ "Non-MPA",
    SITE == "MOHK" ~ "Non-MPA")) %>% 
  filter(SIZE > 82.6) %>% 
  group_by(SITE) %>% 
  summarize(
    LEGAL_LOBSTARS = round(mean(SIZE),2) 
  )

sizes_17_summary

sizes_17_table <- sizes_17_summary %>%
  kable(align = "c", col.names = c("Site", "Percentage")) %>%
  kable_styling(bootstrap_options = c("striped", "consensed"), full_width = F) %>%
  add_header_above(c("Proportion of Legal Lobsters Across Sites in 2017" = 2))

sizes_17_table

#chi-square test to compare the proportions across the five sites
sizes_17_chi <- chisq.test(sizes_17_summary$LEGAL_LOBSTARS)
sizes_17_chi

##Hypothesis Test
#H0: There is no signifigant asscociation between legal carapace size across the five sites.
#HA: There is a signifigant asscociation between legal carapace size across the five sites. 

#X^2 = 0.9995



```

Based on the LTER data of lobster carapace length in 2017 at the different sites (AQUE n = 67, IVEE n = 605, NAPL n = 111, CARP n = 704, MOHK n = 177), carapace length does not differ signifigantly across sites ($\chi^2$(`r sizes_17_chi$parameter`) *p* = `r round(sizes_17_chi$p.value,3)`, $\alpha$ = 0.05). 