---
title: "206 Assignment 4"
author: "Sidney Gerst"
date: "November 13, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load Packages and read in files

```{r}
library(tidyverse)
library(pwr)
library(knitr)
library(kableExtra)
library(plotly)
library(ggrepel)
library(effsize)
library(vcdExtra)
library(dplyr)

lobster_pressure <- read_csv("lobster_traps.csv")

lobster_size <- read_csv("lobster_size_abundance.csv")



```

Clean up Data before converting to tidy

```{r}
size_clean <- lobster_size %>%
  select(YEAR, SITE, SIZE, COUNT) %>%
  filter(COUNT != "0")

size_clean

pressure_clean <- lobster_pressure %>%
  select(YEAR, SITE, TRAPS) 
  
pressure_clean



```


Part 1: Lobster abundance and fishing pressure
- Exploratory data analysis
    - Data frame and scatter plots with 5 locations with counts and fishing pressure (traps) for 2012-2017 (separately)
    - Finalized data summary table for counts/traps at each of the locations by year *** Need to do this
    - Finalized Histogram for each location with count/fishing pressure *** Need to do this
    - Finalized QQ-Plots for each location  *** Need to do this
    
```{r}

abundance <- size_clean %>% 
  select(YEAR, SITE, COUNT) %>% 
  group_by(SITE, YEAR) %>% 
  summarize (
      total_count = round(sum(COUNT), digits = 1)
    ) %>% 
  ggplot(aes(x=YEAR, y = total_count))+
  geom_point() +
  facet_wrap(~SITE, scale = "free") +
  xlab("Year") +
  ylab("Total Count")
    

abundance




pressure <- pressure_clean %>% 
  select(YEAR, SITE, TRAPS) %>%
  filter(SITE == "AQUE"| SITE == "NAPL" | SITE == "MOHK"| SITE == "IVEE"| SITE == "CARP") %>% 
  group_by(SITE, YEAR) %>% 
  summarize (
      total_traps = round(sum(TRAPS), digits = 1)
    ) %>% 
  ggplot(aes(x=YEAR, y = total_traps))+
  geom_point() +
  facet_wrap(~SITE, scale = "free") +
  xlab("Year") +
  ylab("Total Traps")

pressure
 


```





