---
title: "206 Assignment 4"
author: "Sidney Gerst"
date: "November 13, 2018"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load Packages and read in files

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(pwr)
library(knitr)
library(kableExtra)
library(plotly)
library(ggrepel)
library(effsize)
library(vcdExtra)
library(dplyr)

lobster_pressure <- read_csv("lobster_traps.csv")

lobster_size <- read_csv("lobster_size_abundance.csv")



```

Clean up Data before converting to tidy

```{r, echo=FALSE, message=FALSE}
size_clean <- lobster_size %>%
  select(YEAR, SITE, SIZE, COUNT) %>%
  filter(COUNT != "0")

size_clean

pressure_clean <- lobster_pressure %>%
  select(YEAR, SITE, TRAPS) 
  
pressure_clean



```


Part 1: Lobster abundance and fishing pressure
- Exploratory data analysis
    - Scatter plots with 5 locations with counts and fishing pressure (traps) for 2012-2017 (separately)
    - Column graphs with total count/traps by site
    - Finalized line graphs for abundance and pressure
    
    
```{r, echo=FALSE, message=FALSE}

## Make scatter plot for annual abundance by site (2012-2017)
abundance <- size_clean %>% 
  select(YEAR, SITE, COUNT) %>% 
  group_by(SITE, YEAR) %>% 
  summarize (
      total_count = round(sum(COUNT), digits = 1)
    ) %>% 
  ggplot(aes(x=YEAR, y = total_count))+
  geom_point() +
  facet_wrap(~SITE, scale = "free") +
  xlab("Year") +
  ylab("Total Count")+
  theme_classic()+
  ggtitle("Total Lobster Abundance (2012-2017) - TH, EM, SG")
  

abundance


## Make column graph of total abundance by site (2012-2017)
abundance_col <- ggplot(size_clean, aes(x = YEAR, y = COUNT))+
  geom_col()+
  facet_wrap(~SITE, scale = "free")+
  theme_classic()+
  ggtitle("Total Lobster Abundance (2012-2017) - TH, EM, SG")

abundance_col


## Make line graph of total abundance within each site (2012-2017)
abundance_line <- as.data.frame(size_clean) %>% 
  expand.dft(freq="COUNT") %>% 
  count(YEAR, SITE) %>% 
  ggplot(aes(x=YEAR, y = n))+
  geom_line(aes(color=SITE)) +
  xlab("Year") +
  ylab("Total Count")+
  ggtitle("Lobster Abundance at LTER Sites in the Santa Barbara Channel - TH, EM, SG")+
  theme_classic()
  

abundance_line


```


```{r, echo=FALSE, message=FALSE}

# Make a scatter plot for annual lobster pressure for each site (2012-2017)
pressure <- pressure_clean %>% 
  select(YEAR, SITE, TRAPS) %>%
  filter(SITE == "AQUE"| SITE == "NAPL" | SITE == "MOHK"| SITE == "IVEE"| SITE == "CARP") %>% 
  group_by(SITE, YEAR) %>% 
  summarize (
      total_traps = round(sum(TRAPS), digits = 1)
    ) %>% 
  ggplot(aes(x=YEAR, y = total_traps))+
  geom_point() +
  facet_wrap(~SITE, scale = "free") +
  xlab("Year") +
  ylab("Total Pressure")+
  ggtitle("Lobster Abundance (2012-2017) - TH, EM, SG")+
  theme_classic()

pressure


# Make final pressure data frame (with just 5 sites)
pressure_final <- pressure_clean %>% 
  select(YEAR, SITE, TRAPS) %>%
  filter(SITE == "AQUE"| SITE == "NAPL" | SITE == "MOHK"| SITE == "IVEE"| SITE == "CARP")
 

# Make a column graph for total pressure for each site (2012-2017)
pressure_col <- ggplot(pressure_final, aes(x = YEAR, y = TRAPS))+
  geom_col()+
  facet_wrap(~SITE, scale = "free")+
  theme_classic()+
  ggtitle("Total Lobster Fishing Pressure (2012-2017) - TH, EM, SG")

pressure_col


# Make line graph of total pressure within each site (2012-2017)
pressure_line <- as.data.frame(pressure_final) %>% 
  expand.dft(freq="TRAPS") %>% 
  count(YEAR, SITE) %>% 
  ggplot(aes(x=YEAR, y = n))+
  geom_line(aes(color=SITE)) +
  xlab("Year") +
  ylab("Total Pressure")+
  ggtitle("Lobster Fishing Pressure at LTER Sites in Santa Barbara Channel - TH, EM, SG")+
  theme_classic()

pressure_line



```

Discuss trends in lobster abundance (counts) and fishing pressure (buoys):
- In each of the five sites, the total lobster abundance is greater in 2017 than it was in 2012. However, the counts vary greatly over the 5 years. At the AQUE site, counts were much greater over the course of all 5 years than at the other sites. CARP had a dramatic increase in 2017, whereas prior to that the counts were very low. AQUE, CARP, and IVEE all ended 2017 with counts higher than the previous year, while the other three sites experienced decline in their final year. An analysis of lobster pressure partially explains trends in lobster abundance. NAPL and IVEE are both Marine Protected Areas (MPA's) and 


Part 2. Compare mean lobster size by site in 2017

- Convert to tidy format
- Exploratory data analysis 
```{r}

# Convert abundance data frame into tidy format

abundance_tidy <- as.data.frame(size_clean) %>% 
  expand.dft(freq="COUNT")

abundance_tidy

#Create data frame for our 5 locations in 2017, selecting size.
size_means <- abundance_tidy %>% 
  

```



